### 2019-11-11 assignment2/submission2.tex
add assignment2-theory

\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage[a4paper, left=2.5cm, right=2.5cm, top=2.5cm, bottom=4cm]{geometry}
\usepackage[ngerman]{babel}
\usepackage{amsmath, amsthm, amssymb, amstext}
\usepackage{listings}
\usepackage{color}
\usepackage{graphicx}
\usepackage{xparse}
\usepackage{fancyhdr}
\usepackage{algorithmicx}
\usepackage{parskip}
\usepackage[table]{xcolor}
\usepackage{tabularx}
\usepackage{enumerate}
\usepackage{enumitem}
%\usepackage{minted}
\usepackage {tikz}
\usetikzlibrary {positioning}

\pagestyle{fancy}

\rhead{{\newcommand\and\\\getauthors}}
\author{Felix Bühler\\2973410 \and Clemens Lieb\\3130838 \and Steffen Wonner\\2862123 \and Fabian Bühler\\2953320}
\lhead{\textbf\gettitle}
\title{\gettitle}
\chead{\getsubtitle}
\subtitle{\getsubtitle}

\addtolength{\headheight}{2\baselineskip}
\renewcommand{\headrulewidth}{0pt}

\newcommand{\gettitle}{Verteilte Systeme I\\Winter Term 2019/20}
\newcommand{\getsubtitle}{G2T1 – Assignment 2 (theoretical part)}
\newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Steffen Wonner \and Fabian Bühler}
\setlength{\headheight}{53pt}

\begin{document}
\maketitle

\section*{1 Parameter Passing and RMI}
\subsection*{a)}
\begin{enumerate}[label=(\roman*)]
	\item 
	\item 
	\item 
\end{enumerate}

\subsection*{b)}

\subsection*{c)}

\section*{2 Chord System}
\subsection*{a)}
\begin{table}[!ht]
	\centering
	\begin{tabular}{|l|l|}
		\hline
		1 &  \\ \hline
		2 &  \\ \hline
		3 &  \\ \hline
		4 &  \\ \hline
		5 &  \\ \hline
	\end{tabular}
	\caption{4}
\end{table}
\begin{table}[!ht]
	\centering
	\begin{tabular}{|l|l|}
		\hline
		1 &  \\ \hline
		2 &  \\ \hline
		3 &  \\ \hline
		4 &  \\ \hline
		5 &  \\ \hline
	\end{tabular}
	\caption{11}
\end{table}
\begin{table}[!ht]
	\centering
	\begin{tabular}{|l|l|}
		\hline
		1 &  \\ \hline
		2 &  \\ \hline
		3 &  \\ \hline
		4 &  \\ \hline
		5 &  \\ \hline
	\end{tabular}
	\caption{16}
\end{table}
\begin{table}[!ht]
	\centering
	\begin{tabular}{|l|l|}
		\hline
		1 &  \\ \hline
		2 &  \\ \hline
		3 &  \\ \hline
		4 &  \\ \hline
		5 &  \\ \hline
	\end{tabular}
	\caption{19}
\end{table}
\begin{table}[!ht]
	\centering
	\begin{tabular}{|l|l|}
		\hline
		1 &  \\ \hline
		2 &  \\ \hline
		3 &  \\ \hline
		4 &  \\ \hline
		5 &  \\ \hline
	\end{tabular}
	\caption{23}
\end{table}
\begin{table}[!ht]
	\centering
	\begin{tabular}{|l|l|}
		\hline
		1 &  \\ \hline
		2 &  \\ \hline
		3 &  \\ \hline
		4 &  \\ \hline
		5 &  \\ \hline
	\end{tabular}
	\caption{26}
\end{table}
\begin{table}[!ht]
	\centering
	\begin{tabular}{|l|l|}
		\hline
		1 &  \\ \hline
		2 &  \\ \hline
		3 &  \\ \hline
		4 &  \\ \hline
		5 &  \\ \hline
	\end{tabular}
	\caption{31}
\end{table}
\subsection*{b)}
\subsection*{c)}
\subsection*{d)}
\subsection*{e)}
\subsection*{f)}

\section*{3 Name Services}
\subsection*{a)}
\begin{table}[!ht]
	\centering
	\begin{tabular}{|l|l|l|}
		\hline
		& Iterative resolution & Recursive resolution \\ \hline
		1. Lookup &  &  \\ \hline
		2. Lookup &  &  \\ \hline
		3. Lookup &  &  \\ \hline
	\end{tabular}
\end{table}
\subsection*{b)}
\begin{table}[!ht]
	\centering
	\begin{tabular}{|l|l|l|}
		\hline
		& Iterative resolution & Recursive resolution \\ \hline
		1. Lookup &  &  \\ \hline
		2. Lookup &  &  \\ \hline
		3. Lookup &  &  \\ \hline
	\end{tabular}
\end{table}
\subsection*{c)}

\begin{enumerate}[label=(\roman*)]
	\item 
	\item 
	\item 
\end{enumerate}

\end{document}