### 2019-11-14 assignment2/submission2.tex
Fix errors and formatting in submission2

\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage[a4paper, left=2.5cm, right=2.5cm, top=2.5cm, bottom=4cm]{geometry}
\usepackage[ngerman]{babel}
\usepackage{amsmath, amsthm, amssymb, amstext}
\usepackage{listings}
\usepackage{color}
\usepackage{graphicx}
\usepackage{xparse}
\usepackage{fancyhdr}
\usepackage{algorithmicx}
\usepackage{parskip}
\usepackage[table]{xcolor}
\usepackage{tabularx}
\usepackage{enumerate}
\usepackage{enumitem}
%\usepackage{minted}
\usepackage {tikz}
\usetikzlibrary {positioning}

\pagestyle{fancy}

\rhead{{\newcommand\and\\\getauthors}}
\author{Felix Bühler\\2973410 \and Clemens Lieb\\3130838 \and Steffen Wonner\\2862123 \and Fabian Bühler\\2953320}
\lhead{\textbf\gettitle}
\title{\gettitle}
\chead{\getsubtitle}
\subtitle{\getsubtitle}

\addtolength{\headheight}{2\baselineskip}
\renewcommand{\headrulewidth}{0pt}

\newcommand{\gettitle}{Distributed systems I\\Winter Term 2019/20}
\newcommand{\getsubtitle}{G2T1 – Assignment 2 (theoretical part)}
\newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Steffen Wonner \and Fabian Bühler}
\setlength{\headheight}{53pt}

\begin{document}
\maketitle

\section*{1 - Parameter Passing and RMI}
\subsection*{a)}
\begin{enumerate}[label=(\roman*)]
	\item 
	\item 
	\item 
\end{enumerate}

\subsection*{b)}

\subsection*{c)}

\section*{2 - Chord System}
\subsection*{a)}

	\begin{minipage}[t]{.25\linewidth}
        \centering
        Node 4:\\
			\begin{tabular}{|l|l|}
				\hline
				1 & 11 \\ \hline
				2 & 11 \\ \hline
				3 & 11 \\ \hline
				4 & 16 \\ \hline
				5 & 23 \\ \hline
			\end{tabular}
	\end{minipage}%
	\begin{minipage}[t]{.25\linewidth}
        \centering
        Node 11:\\
		\begin{tabular}{|l|l|}
			\hline
			1 & 16 \\ \hline
			2 & 16 \\ \hline
			3 & 16 \\ \hline
			4 & 19 \\ \hline
			5 & 31 \\ \hline
		\end{tabular}
	\end{minipage}%
	\begin{minipage}[t]{.25\linewidth}
        \centering
        Node 16:\\
		\begin{tabular}{|l|l|}
			\hline
			1 & 19 \\ \hline
			2 & 19 \\ \hline
			3 & 23 \\ \hline
			4 & 26 \\ \hline
			5 & 4 \\ \hline
		\end{tabular}
    \end{minipage}
    \begin{minipage}[t]{.25\linewidth}
        \centering
        Node 19:\\
        \begin{tabular}{|l|l|}
            \hline
            1 & 23 \\ \hline
            2 & 23 \\ \hline
            3 & 23 \\ \hline
            4 & 31 \\ \hline
            5 & 4 \\ \hline
        \end{tabular}
    \end{minipage}%

	\begin{minipage}[t]{.25\linewidth}
		\centering
        Node 23:\\
		\begin{tabular}{|l|l|}
			\hline
			1 & 26 \\ \hline
			2 & 26 \\ \hline
			3 & 31 \\ \hline
			4 & 31 \\ \hline
			5 & 11 \\ \hline
		\end{tabular}
	\end{minipage}%
	\begin{minipage}[t]{.25\linewidth}
		\centering
        Node 26:\\
		\begin{tabular}{|l|l|}
			\hline
			1 & 31 \\ \hline
			2 & 31 \\ \hline
			3 & 31 \\ \hline
			4 & 4 \\ \hline
			5 & 11 \\ \hline
		\end{tabular}
	\end{minipage}
	\begin{minipage}[t]{.25\linewidth}
		\centering
        Node 31:\\
		\begin{tabular}{|l|l|}
			\hline
			1 & 4 \\ \hline
			2 & 4 \\ \hline
			3 & 4 \\ \hline
			4 & 11 \\ \hline
			5 & 16 \\ \hline
		\end{tabular}
	\end{minipage}

\subsection*{b)}
\begin{enumerate}
	\item node 4 $ \rightarrow $ node 16
	\item node 16 $ \rightarrow $ node 19
	\item node 19 $ \rightarrow $ node 23
\end{enumerate}

\subsection*{c)}
\begin{minipage}[t]{.25\linewidth}
	\centering
    Node 24:\\
	\begin{tabular}{|l|l|}
		\hline
		1 & 26 \\ \hline
		2 & 26 \\ \hline
		3 & 31 \\ \hline
		4 & 4 \\ \hline
		5 & 11 \\ \hline
	\end{tabular}
\end{minipage}

The finger table of node 16 will change eventually while the one for node 23 will change immediately.

\subsection*{d)}
If both nodes join simultaneously they are between the same two nodes (4 and 11). If they now try to each update their predecessor and successor then we have a race condition while updating a linked list. This can leave the list (or in this case ring) in an inconsistent state that makes routing impossible!

\subsection*{e)}
\subsection*{f)}

\section*{3 - Name Services}

\subsection*{a)}
\begin{center}
	\begin{tabular}{|c|c|c|}
		\hline
		& Iterative resolution & Recursive resolution \\ \hline
		1. Lookup & 10 & 10 \\ \hline
		2. Lookup & 10 & 10 \\ \hline
		3. Lookup & 8 & 8 \\ \hline
	\end{tabular}
\end{center}

\subsection*{b)}
\begin{center}
	\begin{tabular}{|c|c|c|}
		\hline
		& Iterative resolution & Recursive resolution \\ \hline
		1. Lookup & 10 & 10 (8 fast, 2 slow) \\ \hline
		2. Lookup & 10 & 8 (6 fast, 2 slow) \\ \hline
		3. Lookup & 8 & 8 (6 fast, 2 slow) \\ \hline
	\end{tabular}
\end{center}

\subsection*{c)}

\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
\subsubsection*{i.}
\begin{lstlisting}
$ dig NS uni-stuttgart.de
...
;; ANSWER SECTION:
uni-stuttgart.de.	2191	IN	NS	minnehaha.rhrk.uni-kl.de.
uni-stuttgart.de.	2191	IN	NS	dns3.belwue.de.
uni-stuttgart.de.	2191	IN	NS	dns1.belwue.de.
uni-stuttgart.de.	2191	IN	NS	dns0.uni-stuttgart.de.
uni-stuttgart.de.	2191	IN	NS	dns1.uni-stuttgart.edu.
...
\end{lstlisting}
There are 5 server responsible for the DNS entry of 'uni-stuttgart.de':
\begin{itemize}
	\item minnehaha.rhrk.uni-kl.de
	\item dns3.belwue.de
	\item dns1.belwue.de
	\item dns0.uni-stuttgart.de
	\item dns1.uni-stuttgart.edu
\end{itemize}
\subsubsection*{ii.}
\begin{lstlisting}
$ dig -x 129.69.216.249
...
;; ANSWER SECTION:
249.216.69.129.in-addr.arpa. 86397 IN	PTR	ipvslogin.informatik.uni-stuttgart.de.
...
\end{lstlisting}
PTR is usefull for resolving an IP address to a domain or hostname (reverse DNS lookup).\\
The given IP resolves to: 'ipvslogin.informatik.uni-stuttgart.de'
\subsubsection*{iii.}
\begin{lstlisting}
$ dig +trace www.uni-stuttgart.de
...
.			347153	IN	NS	h.root-servers.net.
.			347153	IN	NS	l.root-servers.net.
.			347153	IN	NS	d.root-servers.net.
.			347153	IN	NS	m.root-servers.net.
.			347153	IN	NS	e.root-servers.net.
.			347153	IN	NS	k.root-servers.net.
.			347153	IN	NS	i.root-servers.net.
.			347153	IN	NS	b.root-servers.net.
.			347153	IN	NS	c.root-servers.net.
.			347153	IN	NS	a.root-servers.net.
.			347153	IN	NS	f.root-servers.net.
.			347153	IN	NS	g.root-servers.net.
.			347153	IN	NS	j.root-servers.net.
...
;; Received 553 bytes from 192.168.0.1#53(192.168.0.1) in 28 ms

de.			172800	IN	NS	z.nic.de.
de.			172800	IN	NS	f.nic.de.
de.			172800	IN	NS	s.de.net.
de.			172800	IN	NS	a.nic.de.
de.			172800	IN	NS	l.de.net.
de.			172800	IN	NS	n.de.net.
...
;; Received 802 bytes from 2001:500:12::d0d#53(g.root-servers.net) in 22 ms

uni-stuttgart.de.	86400	IN	NS	dns0.uni-stuttgart.de.
uni-stuttgart.de.	86400	IN	NS	dns1.belwue.de.
uni-stuttgart.de.	86400	IN	NS	dns1.uni-stuttgart.edu.
uni-stuttgart.de.	86400	IN	NS	dns3.belwue.de.
uni-stuttgart.de.	86400	IN	NS	minnehaha.rhrk.uni-kl.de.
...
;; Received 619 bytes from 2001:67c:1011:1::53#53(n.de.net) in 25 ms

www.uni-stuttgart.de.	3600	IN	A	129.69.8.19
...
uni-stuttgart.de.	3600	IN	NS	dns0.uni-stuttgart.de.
uni-stuttgart.de.	3600	IN	NS	minnehaha.rhrk.uni-kl.de.
uni-stuttgart.de.	3600	IN	NS	dns1.uni-stuttgart.edu.
uni-stuttgart.de.	3600	IN	NS	dns3.belwue.de.
uni-stuttgart.de.	3600	IN	NS	dns1.belwue.de.
...
;; Received 1365 bytes from 129.143.2.10#53(dns1.belwue.de) in 25 ms

\end{lstlisting}
The path the lookup took was:
\begin{enumerate}
	\item 192.168.0.1
	\item g.root-servers.net
	\item n.de.net
	\item dns1.belwue.de
\end{enumerate}

\end{document}