### 2019-11-27 assignment3/submission3.tex
Solution for Task 2 added

\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage[a4paper, left=2.5cm, right=2.5cm, top=2.5cm, bottom=4cm]{geometry}
\usepackage[english]{babel}
\usepackage{amsmath, amsthm, amssymb, amstext}
\usepackage{listings}
\usepackage{color}
\usepackage{graphicx}
\usepackage{xparse}
\usepackage{fancyhdr}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{parskip}
\usepackage[table]{xcolor}
\usepackage{tabularx}
\usepackage{enumerate}
\usepackage{enumitem}
%\usepackage{minted}
\usepackage {tikz}
\usetikzlibrary {positioning}

\pagestyle{fancy}

\rhead{{\newcommand\and\\\getauthors}}
\author{Felix Bühler\\2973410 \and Clemens Lieb\\3130838 \and Steffen Wonner\\2862123 \and Fabian Bühler\\2953320}
\lhead{\textbf\gettitle}
\title{\gettitle}
\chead{\getsubtitle}
\subtitle{\getsubtitle}

\addtolength{\headheight}{2\baselineskip}
\renewcommand{\headrulewidth}{0pt}

\newcommand{\gettitle}{Distributed systems I\\Winter Term 2019/20}
\newcommand{\getsubtitle}{G2T1 – Assignment 3 (theoretical part)}
\newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Steffen Wonner \and Fabian Bühler}
\setlength{\headheight}{53pt}

\begin{document}
\maketitle

\section*{1 - Physical Clocks}
\subsection*{a)}

\subsection*{b)}

\section*{2 - Logical Clocks}
\subsection*{a)}
\subsection*{a)}
\begin{enumerate}[label=(\roman*)]
	\item 
	C(e$^1_1$) = 2\\
	C(e$^2_1$) = 3\\
	C(e$^3_1$) = 4\\
	C(e$^4_1$) = 5\\

	C(e$^1_2$) = 5\\
	C(e$^2_2$) = 6\\
	C(e$^3_2$) = 7\\
	C(e$^4_2$) = 8\\

	C(e$^1_3$) = 1\\
	C(e$^2_3$) = 2\\
	C(e$^3_3$) = 3\\
	C(e$^4_3$) = 9\\
	\item 
	VC(e$^1_1$) = \texttt{[1 0 1]}\\
	VC(e$^2_1$) = \texttt{[2 0 1]}\\
	VC(e$^3_1$) = \texttt{[3 0 1]}\\
	VC(e$^4_1$) = \texttt{[4 0 1]}\\

	VC(e$^1_2$) = \texttt{[3 1 1]}\\
	VC(e$^2_2$) = \texttt{[3 2 2]}\\
	VC(e$^3_2$) = \texttt{[3 3 2]}\\
	VC(e$^4_2$) = \texttt{[3 4 2]}\\

	VC(e$^1_3$) = \texttt{[0 0 1]}\\
	VC(e$^2_3$) = \texttt{[0 0 2]}\\
	VC(e$^3_3$) = \texttt{[0 0 3]}\\
	VC(e$^4_3$) = \texttt{[3 4 4]}\\
	\item
	P$_1$: e$^1_1$ e$^2_1$ e$^3_1$\\
	P$_2$: e$^1_2$ e$^2_2$\\
	P$_3$: e$^1_3$ e$^2_3$\\
	The vector event time of those events is smaller than the vector event time of e$^3_2$. A smaller vector time means events are causally related.
\end{enumerate}
\subsection*{b)}
\begin{enumerate}[label=(\roman*)]
	\item 
	P$_1$: e$_6$ e$_5$ e$_9$ e$_4$ e$_8$ e$_{10}$\\
	P$_2$: e$_{12}$ e$_3$ e$_{13}$\\
	P$_3$: e$_{11}$ e$_7$ e$_1$ e$_2$\\
	\item 
	Local event: e$_5$ e$_9$ e$_3$ e$_1$ e$_{11}$\\
	Send event: e$_6$ e$_4$ e$_{10}$ e$_{12}$\\
	Recieve event:  e$_8$ e$_{13}$ e$_7$ e$_2$\\
\end{enumerate}
\section*{3 - Global State}
\subsection*{a)}

\subsection*{b)}

\subsection*{c)}

\section*{4 - Snapshot Algorithm}
\subsection*{a)}

\subsection*{b)}

\end{document}