### 2019-12-10 assignment4/submission4.tex
done 4.1 c) and d)

\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage[a4paper, left=2.5cm, right=2.5cm, top=2.5cm, bottom=4cm]{geometry}
\usepackage[english]{babel}
\usepackage{amsmath, amsthm, amssymb, amstext}
\usepackage{listings}
\usepackage{color}
\usepackage{graphicx}
\usepackage{xparse}
\usepackage{fancyhdr}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{parskip}
\usepackage[table]{xcolor}
\usepackage{tabularx}
\usepackage{enumerate}
\usepackage{enumitem}
%\usepackage{minted}
\usepackage {tikz}
\usetikzlibrary{positioning}

\pagestyle{fancy}

\rhead{{\newcommand\and\\\getauthors}}
\author{Felix Bühler\\2973410 \and Clemens Lieb\\3130838 \and Steffen Wonner\\2862123 \and Fabian Bühler\\2953320}
\lhead{\textbf\gettitle}
\title{\gettitle}
\chead{\getsubtitle}
\subtitle{\getsubtitle}

\addtolength{\headheight}{2\baselineskip}
\renewcommand{\headrulewidth}{0pt}

\newcommand{\gettitle}{Distributed systems I\\Winter Term 2019/20}
\newcommand{\getsubtitle}{G2T1 – Assignment 4 (theoretical part)}
\newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Steffen Wonner \and Fabian Bühler}
\setlength{\headheight}{53pt}

\begin{document}
\maketitle

\section*{1 - Global State}
\subsection*{a)}
\subsection*{b)}
\subsection*{c)}

\tikzstyle{nofill}=[rectangle,draw, align=center]
\tikzstyle{green}=[rectangle,draw,fill=green!50,align=center]
\tikzstyle{blue}=[rectangle,draw,fill=blue!50,align=center]
\tikzstyle{both}=[rectangle,draw,fill=red!50!,align=center]

\begin{figure}[!ht]
\centering
\begin{tikzpicture}[auto,node distance=1.5cm]
\node[green] (s00) {$ S_{00} $ \\ $ x_1 = 0 $ \\ $ x_2 = 0 $};

\node[both] (s10) [below left = of s00] {$ S_{10} $ \\ $ x_1 = 10 $ \\ $ x_2 = 0 $};

\node[both] (s20) [below left = of s10] {$ S_{20} $ \\ $ x_1 = 5 $ \\ $ x_2 = 0 $};
\node[green] (s11) [below right = of s10] {$ S_{11} $ \\ $ x_1 = 10 $ \\ $ x_2 = 20 $};

\node[both] (s30) [below left = of s20] {$ S_{30} $ \\ $ x_1 = 15 $ \\ $ x_2 = 0 $};
\node[green] (s21) [below left = of s11] {$ S_{21} $ \\ $ x_1 = 5 $ \\ $ x_2 = 20 $};
\node[green] (s12) [below right = of s11] {$ S_{12} $ \\ $ x_1 = 10 $ \\ $ x_2 = 30 $};

\node[nofill] (s31) [below left = of s21] {$ S_{31} $ \\ $ x_1 = 15 $ \\ $ x_2 = 20 $};
\node[green] (s22) [below left = of s12] {$ S_{22} $ \\ $ x_1 = 5 $ \\ $ x_2 = 30 $};

\node[green] (s32) [below left = of s22] {$ S_{32} $ \\ $ x_1 = 15 $ \\ $ x_2 = 30 $};

\node[both] (s42) [below left = of s32] {$ S_{42} $ \\ $ x_1 = 55 $ \\ $ x_2 = 30 $};

\node[both] (s52) [below left = of s42] {$ S_{52} $ \\ $ x_1 = 60 $ \\ $ x_2 = 30 $};

\node[both] (s53) [below right = of s52] {$ S_{53} $ \\ $ x_1 = 60 $ \\ $ x_2 = 13 $};

\path (s10) edge node {} (s00);

\path (s10) edge node {} (s11);
\path (s10) edge node {} (s20);

\path (s11) edge node {} (s21);
\path (s20) edge node {} (s21);
\path (s20) edge node {} (s30);
\path (s11) edge node {} (s12);

\path (s30) edge node {} (s31);
\path (s21) edge node {} (s31);
\path (s21) edge node {} (s22);
\path (s12) edge node {} (s22);

\path (s31) edge node {} (s32);
\path (s22) edge node {} (s32);

\path (s32) edge node {} (s42);

\path (s42) edge node {} (s52);

\path (s52) edge node {} (s53);
\end{tikzpicture}
\caption{solution for 1.c)}
\end{figure}

\subsection*{d)}
\begin{itemize}
	\item $ \phi_{1} = (x_{2} - x_{1}) = 5 $\\
	$ \Rightarrow (\neg\phi_{1}) $ marked in \textcolor{blue}{blue}
	\item $ \phi_{2} = (x_{2} - x_{1}) \leq 5 $ and $ (x_{2} - x_{1}) > 0 $\\
	$ \Rightarrow (\phi_{2}) $ marked in \textcolor{green}{green}
	\item if they fulfill $ (\neg\phi_{1}) $ and $ (\phi_{2}) $ marked in \textcolor{red}{red}
\end{itemize}

\begin{enumerate}
	\item safety condition $ \neg\phi_{1} $ is not fulfilled in the state $ S_{31} $and thereby not fulfilled.
	\item liveness condition $ \phi_{2} $ is fulfilled in the end and thereby fulfilled.
\end{enumerate}
\section*{2 - Transaction Processing}
\subsection*{a)}
\subsection*{b)}

\section*{3 - Two-Phase Locking}

\end{document}